---
title: "MIS 64018 - Assignment 4"
author: "Tejasvini Mavuleti"
date: "2022-10-24"
output: word_document
---
# Hope Valley Health Care Association

Inputs -  Staffing Labor, Cost of Supplies
Outputs -  No of patient-days reimbursed by third party, No of patient-days reimbursed privately 

```{r}
# First we start off by calling the Benchmarking package to run the DEA analysis. 
require(Benchmarking)
```

Building a matrix to load the data 
```{r}
# Add the input variables to the matrix
X <- matrix(c(150, 400, 320, 520, 350, 320, 0.2, 0.7, 1.2, 2.0, 1.2, 0.7), ncol = 2)

# Add the output variables to the matrix
Y <- matrix(c(14000, 14000, 42000, 28000, 19000, 14000, 3500, 21000, 10500, 42000, 25000, 15000), ncol = 2)

# Name the columns as they were in the data set
colnames(X) <- c("Staff Hours per Day","Supplies per Day")
colnames(Y) <- c("Reimburse Patient-Days", "Privately Paid Patient-Days")

# Review the data set in the matrix form
print(X)
print(Y)
```
# Solving Q1 and Q2 together 
# Q1. - Formuulate and perform DEA analysis
# Q2. - Determine the Peers and Lambdas under each of the above assumptions 

```{r}
# Formulating DEA analysis using FDH

FDH <- rep("FDH", times = 6)
Not_Applicable <- rep(NA, times = 6)
DEA_FDH <- dea(X, Y, RTS = "FDH")
DEA_FDH_Peers <- peers(DEA_FDH)                             
DEA_FDH_Lambda <- lambda(DEA_FDH)  

print(DEA_FDH)
print(DEA_FDH_Peers)
print(DEA_FDH_Lambda)

dea.plot.isoquant(X, Y, RTS= "FDH")

# Plot and summarize the results for the FDH assumption 

DEA_FDH_Peers <- cbind(DEA_FDH_Peers, Not_Applicable, Not_Applicable)
FDH_Summary <- cbind(FDH, DEA_FDH$eff, DEA_FDH_Peers, DEA_FDH_Lambda)
colnames(FDH_Summary) <- c("Method","Eff", "P1", "P2", "P3", "L1", "L2", "L3", "L4", "L5", "L6")
print(FDH_Summary)
```



```{r}
# Formulating DEA analysis using CRS
CRS <- rep("CRS", times = 6)
DEA_CRS <- dea(X, Y, RTS = "CRS")
DEA_CRS_Peers <- peers(DEA_CRS)                             
DEA_CRS_Lambda <- lambda(DEA_CRS)

print(DEA_CRS)
print(DEA_CRS_Peers)
print(DEA_CRS_Lambda)

dea.plot.isoquant(X, Y, RTS= "CRS")

#Plot and summarize the results for the CRS assumption 

DEA_CRS_Lambda <- cbind(DEA_CRS_Lambda, Not_Applicable, Not_Applicable)
CRS_Summary <- cbind(CRS, DEA_CRS$eff, DEA_CRS_Peers, DEA_CRS_Lambda)
colnames(CRS_Summary) <- c("Method","Eff", "P1", "P2", "P3", "L1", "L2", "L3", "L4", "L5", "L6")
CRS_Summary <- as.data.frame(CRS_Summary)
CRS_Summary
```


```{r}
# Formulating DEA analysis using VRS
VRS <- rep("VRS", times = 6)
DEA_VRS <- dea(X, Y, RTS = "VRS")
DEA_VRS_Peers <- peers(DEA_VRS)                             
DEA_VRS_Lambda <- lambda(DEA_VRS)

print(DEA_VRS)
print(DEA_VRS_Peers)
print(DEA_VRS_Lambda)

dea.plot.isoquant(X, Y, RTS= "VRS")

# Plot and summarize the results for the VRS assumption 

DEA_VRS_Lambda <- cbind(DEA_VRS_Lambda, Not_Applicable)
VRS_Summary <- cbind(VRS, DEA_VRS$eff, DEA_VRS_Peers, DEA_VRS_Lambda)
colnames(VRS_Summary) <- c("Method","Eff", "P1", "P2", "P3", "L1", "L2", "L3", "L4", "L5", "L6")
VRS_Summary <- as.data.frame(VRS_Summary)
VRS_Summary
```


```{r}
# Formulating DEA analysis using IRS
IRS <- rep("IRS", times = 6)
DEA_IRS <- dea(X, Y, RTS = "IRS")
DEA_IRS_Peers <- peers(DEA_IRS)                             
DEA_IRS_Lambda <- lambda(DEA_IRS) 

print(DEA_IRS)
print(DEA_IRS_Peers)
print(DEA_IRS_Lambda)

dea.plot.isoquant(X, Y, RTS= "IRS")

# Plot and summarize the results for the IRS assumption

DEA_IRS_Lambda <- cbind(DEA_IRS_Lambda, Not_Applicable)
IRS_Summary <- cbind(IRS, DEA_IRS$eff, DEA_IRS_Peers, DEA_IRS_Lambda)
colnames(IRS_Summary) <- c("Method","Eff", "P1", "P2", "P3", "L1", "L2", "L3", "L4", "L5", "L6")
IRS_Summary <- as.data.frame(IRS_Summary)
IRS_Summary
```


```{r}
# Formulating DEA analysis using DRS
DRS <- rep("DRS", times = 6)
DEA_DRS <- dea(X, Y, RTS = "DRS")
DEA_DRS_Peers <- peers(DEA_DRS)                             
DEA_DRS_Lambda <- lambda(DEA_DRS)

print(DEA_DRS)
print(DEA_DRS_Peers)
print(DEA_DRS_Lambda)

dea.plot.isoquant(X, Y, RTS= "DRS")

# Plot and summarize the results for the DRS assumption

DEA_DRS_Lambda <- cbind(DEA_DRS_Lambda, Not_Applicable, Not_Applicable)
DRS_Summary <- cbind(DRS, DEA_DRS$eff, DEA_DRS_Peers, DEA_DRS_Lambda)
colnames(DRS_Summary) <- c("Method","Eff", "P1", "P2", "P3", "L1", "L2", "L3", "L4", "L5", "L6")
DRS_Summary <- as.data.frame(DRS_Summary)
DRS_Summary
```

```{r}
# Formulating DEA analysis using FRH or also called as ADD
ADD <- rep("ADD", times = 6)
DEA_ADD <- dea(X, Y, RTS = "ADD")
DEA_ADD_Peers <- peers(DEA_ADD)                             
DEA_ADD_Lambda <- lambda(DEA_ADD)

print(DEA_ADD)
print(DEA_ADD_Peers)
print(DEA_ADD_Lambda)

# Plot and summarize the results for the FRH/ADD assumption

DEA_ADD_Peers <- cbind(DEA_ADD_Peers, Not_Applicable, Not_Applicable)
ADD_Summary <- cbind(ADD, DEA_ADD$eff, DEA_ADD_Peers, DEA_ADD_Lambda)
colnames(ADD_Summary) <- c("Method","Eff", "P1", "P2", "P3", "L1", "L2", "L3", "L4", "L5", "L6")
ADD_Summary <- as.data.frame(ADD_Summary)
ADD_Summary
```

# Q3. Summarize your results in a tabular format 
```{r}
# Summarizing all the results from every DEA assumption in a tabular form 
summary_table <- rbind(FDH_Summary, CRS_Summary, VRS_Summary, IRS_Summary, DRS_Summary, ADD_Summary)
print(summary_table)

# creating an excel file
library(rio)
library(writexl)
export(summary_table,"C:/Users/mavul/OneDrive/Documents/MIS 64018 - Assignment 4/64018-A4.xlsx")
```

# Q4. Compare and contrast the above results 
In conclusion - 
All the DEA assumptions estimate the technology by using a minimum extrapolation technique. What we learn here, is that the FDH is the smallest technology set and it gives out fewer outputs (no. of patient days reimbursed by third-party sources and no. of patient days reimbursed privately) with more inputs (no. of patient days reimbursed by third-party sources and no. of patient days reimbursed privately) (staffing labor and the cost of supplies). 

Key takeaways - 
- From the above results we saw that the FRH and FDH assumptions gave out efficiency with the value 1.0. And in every DEA assumptions the peer and lambda values were identical. 
- The other observation we found out was about the efficiency at 1.0 - 
In the CRS method got DMU[1:4] efficient at 1.0
VRS method had DMU[1:5] efficient at 1.0
IRS method had DMU[1:5] efficient at 1.0
DRS method had DMU[1:4] efficient at 1.0
- Most importantly all of the other less efficient DMUs had a Peer[1] and Peer [2] value of 1 and 2, but Peer[3] had a value of either 4 or 5. This was dependent on the method we used for the DEA assumption. 
- The lambdas also known as relative weights for the same DMUs were extremely close to each other in every method.

What we know about every Facility - 
- Facilities 1,2,3 and 4 are fully efficient for all the assumptions and Facilities 5 and 6 are not efficient.
- Facility 5 is fully efficient for FDH,VRS,IRS and FRH assumptions.
- We can see that there is 97.7% efficiency for the CRS and DRS assumptions.
- Facility 6 is fully efficient for the FDH and FRS assumptions.
- For Facility 6, CRS and DRS assumptions are 86.7% efficient
- Lastly, for Facility 6, IRS and VRS assumptions are 89.6% efficient. 



